---
title: "Darwin Core Mapping"
subtitle: "Bullfrog DWC mapping"
author: "Dimitri Brosens"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

# Setup 

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = TRUE)
```

Load libraries:

```{r message = FALSE}
library(dplyr)      # To do data science
library(tidylog)        # To provide feedback on dplyr functions
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
#library(rgbif)          # To use GBIF services
#library(uuid)           # To generate UUID's
library(readr)
library(stringr)
library(tidyr)
```

# Read source data



```{r}
input_data <- read_delim(here::here("output", "gbif_stierkikker_vangsten.csv"), ",")


```


Preview data:

```{r}
input_data %>% head()
```



# Preprocessing

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% 
  remove_empty("rows") %<>%
  clean_names()
```




### fix unique ID issue

```{r}
input_data <- mutate_all(input_data, as.character())
```


# Darwin Core mapping

Map the data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml).

```{r}
occurrence <- input_data
```

### type

```{r}
occurrence %<>% mutate(dwc_type = "Event")
```

### language

```{r}
occurrence %<>% mutate(dwc_language = "en")
```

### license

```{r}
occurrence %<>% mutate(dwc_license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

### rightsHolder

```{r}
occurrence %<>% mutate(dwc_rightsHolder = "INBO")
```

### accessRights

```{r}
occurrence %<>% mutate(dwc_accessRights = "http://www.inbo.be/en/norms-for-data-use") 
```

### datasetID

```{r}
occurrence %<>% mutate(dwc_datasetID = "https://doi.org/10.15468/daf62d")
```

### institutionCode


### datasetName

```{r}
occurrence %<>% mutate(dwc_datasetName = "Invasive species - American bullfrog (Lithobates catesbeianus) in Flanders, Belgium (post 2018)")
```

### basisOfRecord

```{r}
occurrence %<>% mutate(dwc_basisOfRecord = "HumanObservation")
```

### occurrenceID


```{r}
occurrence %<>% mutate(dwc_occurrenceID = occurrence_id)
```




### recordedBy

```{r}
occurrence %<>% mutate(dwc_recordedBy =  recorded_by)
```

### individualCount

```{r}
#mutate(occurrence$mm)as.numeric()
  
occurrence %<>% mutate(dwc_individualCount = individual_count)

```




### occurrenceStatus

```{r}
occurrence %<>% mutate(dwc_occurrenceStatus = case_when(
  dwc_individualCount > 0 ~ "present",
  dwc_individualCount == 0 ~ "absent"
))
```

```{r}
occurrence %<>% mutate(dwc_occurrenceRemarks = 'target species')
```



### eventID


```{r}
occurrence %<>% mutate(dwc_eventID = str_replace_all(event_id, " ",""))
```

### eventDate

```{r}
occurrence %<>% mutate(dwc_eventDate = datum)
```




### habitat

```{r}
occurrence %<>% mutate(dwc_habitat = recode(type
                                            , 'Stilstaand' = 'freshwater; stagnant'))
```

### samplingProtocol

```{r}
occurrence %<>% mutate(dwc_samplingProtocol = recode(actie
                                                     ,"Afvangst" = "catch, fike"
                                                     ,"YPT" = "yellow pan tras"
                                                     ,"MT" = "Malaise trap"
                                                     ,"RsPT" = "pink pan traps"
                                                     ,"HC" = "collected by hand"
                                                     ,"YPT/RPT" = "yellow and red pan traps"
                                                     ,"WPT" = "white pan traps"
                                                     ,"BPT" = "blue pan traps"
                                                     )) 
```
### sampleSizeValue

```{r}
occurrence %<>% mutate(dwc_sampleSizeValue = no_fuiken) 
```
### sampleSizeUnit

```{r}
occurrence %<>% mutate(dwc_sampleSizeUnit = recode(fuiktype
                                                   ,"dubbele fuik" = "paired fyke"
                                                   ,"salamanderfuik drijvend NA" = "amphibian trap floating"
                                                   ,"salamanderfuik bodem" = "amphibian trap bottom"
                                                   ,"enkele fuik" = "single fyke"
                                                   ,"vrijwilligers Natuurpunt" = "unspecified fyke net"
                                                   , .default = "fyke"
                                                     )) 
```


### sex

```{r}
occurrence %<>% mutate (dwc_sex = recode(lifestage 
                                               ,"M2" = "unknown"
                                               ,"AM" = "male"
                                               ,"AV" = "female"
                                               ,"L0" = "unknown"
                                               ,"L1" = "unknown"
                                               ,"L00" = "unknown"
                                               ,"L2" = "unknown"
                                               ,"M1" = "unknown"))

```

### lifeStage

```{r}
occurrence %<>% mutate (dwc_lifeStage = recode(lifestage 
                                                ,"M2" = "metamorph without tail"
                                               ,"AM" = "adult"
                                               ,"AV" = "adult"
                                               ,"L0" = "larvae no legs"
                                               ,"L1" = "larvae with hind legs < 1cm"
                                               ,"L00" = "smallest larvae < 5cm no legs"
                                               ,"L2" = "larvae with hind legs > 1cm"
                                               ,"M1" = "metamorph with tail"))



```


### locationID

```{r}
occurrence %<>% mutate(dwc_locationID = location_id)
```


### locality

```{r}
occurrence %<>% mutate(dwc_locality = paste(locatie))

```


### continent

```{r}
occurrence %<>% mutate(dwc_continent = "Europe")
```

### countryCode

```{r}
occurrence %<>% mutate(dwc_countryCode = 'BE')
```



### locationRemarks



### decimalLatitude

```{r}
occurrence %<>% mutate(dwc_decimalLatitude = sprintf(
  "%.5f",
  round(lat, digits = 5)
))
```

### decimalLongitude

```{r}
occurrence %<>% mutate(dwc_decimalLongitude = sprintf(
  "%.5f",
  round(long, digits = 5)
))
```

### geodeticDatum

```{r}
occurrence %<>% mutate(dwc_geodeticDatum = "WGS84") 
```

### coordinateUncertaintyInMeters

```{r}
occurrence %<>% mutate(dwc_coordinateUncertaintyInMeters = sqrt(2*500^2))
```
### Identification


```{r}
occurrence %<>% mutate(dwc_identifiedBy = identified_by)
```

```{r}
#occurrence %<>% mutate(dwc_identifiedByID = "https://orcid.org/0000-0001-5198-5928")
```


```{r}
#occurrence %<>% mutate(dwc_dateIdentified = identified_date)
```

```{r}
occurrence %<>% mutate(dwc_identificationVerificationStatus = "validated")
```


### scientificName

```{r}
occurrence %<>% mutate(dwc_scientificName = scientificname)
```

### kingdom

```{r}
occurrence %<>% mutate(dwc_kingdom = "Animalia")
```

### vernacularName


### taxonRank

```{r}
occurrence %<>% mutate(dwc_taxonRank = "species")
```

### nomenclaturalCode

```{r}
occurrence %<>% mutate(dwc_nomenclaturalCode = "ICZN")
```

## Post-processing

```{r}
occurrence %<>% select(starts_with("dwc_"))
colnames(occurrence) <- str_remove(colnames(occurrence), "dwc_")
```

```{r}
occurrence <- mutate_all(occurrence, as.character())
```

```{r}
occurrence$eventDate <- str_replace_all(occurrence$eventDate, "NA", "")
```


```{r}
#occurrence <-  mutate_at(c(occurrence$eventDate), ~replace_na(.,""))

```



```{r}
#occurrence %<>% distinct(occurrenceID2, .keep_all = TRUE)
```

Preview data:

```{r}
occurrence %>% head()
```

Save to CSV:

```{r}
write_csv(occurrence, here::here("darwincore", "processed", "occurrenceSK.csv"), na = "")
```

# Checks
# Checks

```{r}
test <- length(unique(occurrence$eventID))
test2 <- length(unique(occurrence$occurrenceID))
n_occur <- data.frame(table(occurrence$occurrenceID))
traps <- distinct(input_data, fuiktype, .keep_all = FALSE)
traps2 <- distinct(occurrence, sampleSizeUnit, .keep_all = FALSE)

```

